---
title: "Predict Horsepower from MPG - 20170429 R9-w4 proj"
author: "Steve Jin"
date: "April 29, 2017"
output:  slidy_presentation #ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Shiny Presentation


The application is based on course lecture, with addtional function added. 

* Model 1 is original model from lecture, which predict housepower based on *mtcars* data.
* Model 2 are modified to predict housepower on modified model.   

In the server program, the brushed point are collected and removed from the mtcars data set and generate new model, which is used for prediction.

* Under plot panel, the brushed data point features are listed in tableOutput format.   

This function creates an alternative function - explorary of the data features - all the features are display.  

* Standalone Applcation URL: https://forest416.shinyapps.io/predict_horsepower_from_mpg_-_20170429_r9-w4_project/

* Presentation URL: https://forest416.shinyapps.io/proj_shiny/

## server side code display
```{r eval=F,echo=TRUE}
server <- function(input, output) {
        model1 <-lm (hp ~ mpg, data=mtcars)
        model2 <-reactive ( {
                mtcars.names <- names(mtcars)
                mtcars.subset <- brushedPoints(mtcars, input$brush1, yvar='hp', xvar='mpg')
                if (dim(mtcars.subset)[1] != 0) {
                        #mtcars.remain <- setDT(mtcars)[!mtcars.subset, key=mtcars.names]
                        dt1 <- data.table(mtcars, key=mtcars.names)
                        dt2 <- data.table(mtcars.subset)
                        mtcars.remain <- dt1[!dt2]
                        lm(hp ~ mpg, data = mtcars.remain)
                } else  { NULL }
        model1pred <- reactive( {
                predict(model1, newdata= data.frame(mpg=input$sliderMPG))
        })
        model2pred <- reactive( {
                if (is.null(model2())) {NULL}
                else {
                        predict(model2(), newdata=data.frame(mpg=input$sliderMPG))   
                }
        })
```

## ui side code display
```{r eval=FALSE, echo=TRUE}
ui = fluidPage(
tags$head( tags$style(HTML( 'body { background-color: #e1f2eb } html {font-size:8t} '))),
  titlePanel('Predict Horsepower from MPG - 20170429 R9-w4 proj', windowTitle = 'Predict Horsepower from MPG'),
  column(  width=12,    
    fluidRow(
        column(4, 
            sliderInput('sliderMPG', 'Select MPG for HP predict: ' , 10, 35, value=20),
            p(strong('Model 1: '),'What\'s the predict HP by given MPG?',
                HTML('<font color="red">'), strong(textOutput("pred1", inline=T)),
                HTML('</font>')),
            checkboxInput('showModel1', 'Show/Hide prediction line. No data excluded', value=T),
            p(strong('Model 2: '),'And What if exclude certain point(s) from predction modeling?', 
                HTML('<font color="blue">'), strong(textOutput("pred2", inline=T)),
                HTML('</font> * <span style="font-size:0.8em">
                Use brush to select points to be excluded for prediction model</span><p/>')),
            checkboxInput('showModel2', 'Show/Hide prediction line.',  value=T)
            ),
        column(8, plotOutput('distPlot', brush = brushOpts(id='brush1')))
    ),
  fluidRow(p('Excluded ',strong(textOutput('text2', inline=T)), 'data point:'),
        column(12, tableOutput('table') )
)))

```


## App demo

```{r, echo=FALSE}
suppressPackageStartupMessages(library(shiny))
suppressPackageStartupMessages(library(data.table))

shinyApp (
options = list(height = 1000, width=800),
server <- function(input, output) {
        model1 <-lm (hp ~ mpg, data=mtcars)
        
        model2 <-reactive ( {
                mtcars.names <- names(mtcars)
                mtcars.subset <- brushedPoints(mtcars, input$brush1, yvar='hp', xvar='mpg')
                if (dim(mtcars.subset)[1] != 0) {
                        #mtcars.remain <- setDT(mtcars)[!mtcars.subset, key=mtcars.names]
                        dt1 <- data.table(mtcars, key=mtcars.names)
                        dt2 <- data.table(mtcars.subset)
                        mtcars.remain <- dt1[!dt2]
                        lm(hp ~ mpg, data = mtcars.remain)
                } else  {
                        NULL
                }
                
                # dataframe minus operation
                # Credit: http://stackoverflow.com/questions/16143445/minus-operation-of-data-frames
                # dt1 <- data.table(df1, key=c('c1', 'c2',...)
                # dt2 <- data.table(df2)
                # dt1[!dt2]
                # 
                # or one-line cmd
                # setDT(df1[!df2, on='c1'])
                # manual:
                # setDT(x, keep.rownames=FALSE, key=NULL, check.names=FALSE)
                
                
        })
        # dt1 <- data.table(df1, key=c('c1', 'c2',...)
        # dt2 <- data.table(df2)
        # dt1[!dt2]
        
        
        
        
        model1pred <- reactive( {
                predict(model1, newdata= data.frame(mpg=input$sliderMPG))
        })
        model2pred <- reactive( {
                if (is.null(model2())) {NULL}
                else {
                        predict(model2(), newdata=data.frame(mpg=input$sliderMPG))   
                }
                
        })
        

output$distPlot <- renderPlot({
        mpgInput <-input$sliderMPG
        par(5.1,4.1,2,2)
        plot(mtcars$mpg, mtcars$hp, xlab = 'Miles per Gallon',
                       ylab='house power', bty = 'n', col=levels(as.factor(mtcars$cyl)), pch=16, cex=1)
        legend('topright', legend=paste('cylinder: ',unique(mtcars$cyl)), col=levels(as.factor(mtcars$cyl)), pch=16)
        
        #plot_ly(mtcars, x=~mpg, y=~hp, xlab = 'Miles per Gallon',
        # p <- ggplot(mtcars, aes(x=mpg, y=hp)) +
        #         geom_point(aes(mpg,hp)) +
        #         geom_jetter(aes(mpg, hp),color = 'green') +
        #         #geom_smooth(aes(mpg,hp), method = lm, color='red',se=F)
        #         geom_line(data=prediectv)
        # ggplotly(p) %>% layout(dragmode = 'select')
        
        if(input$showModel1) {
                abline(model1, col = 'red', lwd=2)
        }
        if(input$showModel2 ) {
                if (!is.null(model2())) {
                        abline(model2(), col = rgb(0,1,1, alpha=.4), lwd=2)
                        }
        }

        points(mpgInput, model1pred(), col = 'red', pch = 16, cex=2)
        points(mpgInput, model2pred(), col = 'blue', pch=16, cex=2)
  })
  output$pred1 <- renderText(model1pred())
  output$pred2 <- renderText(model2pred())
 
   xystr<-function(x){
          xname<-names(x)[1]
          yname<-names(x)[2]
          lines<-apply(x,1, function(v) paste0(xname, "=", v[1], ", ", yname,"=",v[2], '\n' ))
          lines.txt <- paste0(unlist(lines))
          lines.txt
                       
  }

  output$table <- renderTable((brushedPoints(mtcars, input$brush1, yvar='hp', xvar='mpg')))
  
  output$brushText <- renderText(xystr(brushedPoints(mtcars, input$brush1, yvar='hp', xvar='mpg')))
  
  output$text2 <- renderText(dim(brushedPoints(mtcars, input$brush1, yvar='hp', xvar='mpg'))[1])
  
 
  
},

ui = fluidPage(
        
# customize html page apperance
tags$head( 
        tags$style(
                HTML( #Arial,Helvetica
                        'body {
                                background-color: #e1f2eb
                        }
                        html {font-size:8t}
                        html {font-family: "Times New Roman", Georgia, Serif}
                        <basefont color="#FF0000" face="Helvetica" size="-3 />
                       '
                )
        )
),
  # Application title
  titlePanel('Predict Horsepower from MPG - 20170429 R9-w4 proj', windowTitle = 'Predict Horsepower from MPG'),
 
  column(  width=12,    
        fluidRow(
                column(4, 
                        sliderInput('sliderMPG', 'Select MPG for HP predict: ' , 10, 35, value=20),
                        p(strong('Model 1: '),'What\'s the predict HP by given MPG?',
                                HTML('<font color="red">'),
                                        strong(textOutput("pred1", inline=T)),
                                HTML('</font>')),
                        checkboxInput('showModel1', 'Show/Hide prediction line. No data excluded', value=T),
                       
                        p(strong('Model 2: '),'And What if exclude certain point(s) from predction modeling?', 
                                HTML('<font color="blue">'),
                                        strong(textOutput("pred2", inline=T)),
                                HTML('</font>
                                     * <span style="font-size:0.8em">
                            Use brush to select points to be excluded for prediction model</span><p/>')),
                        checkboxInput('showModel2', 'Show/Hide prediction line.',  value=T)
                        
                       
                ),
                column(8,
                       plotOutput('distPlot', brush = brushOpts(id='brush1'))
                      
                )
        ),
  fluidRow(
          p('Excluded ',strong(textOutput('text2', inline=T)), 'data point:'),
                #column(2, dataTableOutput('table') )
                column(12, tableOutput('table') )
        )
      )
)
)
```






